---
title: "ST 558 Project 2"
author: "Kaylee Frazier and Rebecca Voelker"
date: "10/31/2021"
output: html_document
---
```{r}
#read in library
library(tidyverse)
```


```{r data, message = FALSE}

getwd()

## Read in Raw Data Using Relative Path
library(readr)
rawData <- read_csv(file = "./project2_rawdata.csv") 
rawData

## Subset Data for Entertainment Data Channel
library(dplyr)
entertainmentData <- rawData %>% filter( data_channel_is_entertainment == 1)

## Create Training and Test Data Sets
library(caret)
set.seed(500)
trainIndex <- createDataPartition(entertainmentData$shares, p = 0.7, list = FALSE)

trainData <- entertainmentData[trainIndex,]
testData <- entertainmentData[-trainIndex,]

## Printed Out Training and Test Data Sets - Can Check if you are getting matching first ~5 Observations (Using Seed = 500)
trainData
testData

```
Create new variable day_of_the_week
```{r}
#mutate new variable using weekday_is_() variables
trainDataNew <- trainData %>% mutate(day_of_the_week =   if_else(weekday_is_monday == 1, "Monday",
       if_else(weekday_is_tuesday == 1, "Tuesday",
               if_else(weekday_is_wednesday == 1, "Wednesday",
                      if_else(weekday_is_thursday == 1, "Thursday",
                              if_else(weekday_is_friday == 1, "Friday",
                                      if_else(weekday_is_saturday == 1, "Saturday", "Sunday"
                                              ))))))) 
trainDataNew
```


Numerical summary of shares by day_of_the_week
```{r}
#find mean, median, and standard deviation of shares by day_of_the_week
daySummary <-trainDataNew %>% group_by(day_of_the_week) %>% summarise(avg_shares = mean(shares), med_shares = median(shares), sd_shares = sd(shares))

#print out numerical summary
daySummary
```
In the summary plot you can see which days had the highest and lowest average shares, median shares, and look at the standard deviation to see which day had the largest range of shares.

Summary table of shares by number of images
```{r}
#create summary table with different ranges of images
imagesSummary <- trainDataNew %>% group_by(cut(num_imgs, c(min(num_imgs), 10, 20, 40, 60, 80, max(num_imgs)))) %>% summarise(avg_shares = mean(shares), med_shares = median(shares), sd_shares = sd(shares))

#rename column name
names(imagesSummary)[names(imagesSummary) == "cut(num_imgs, c(min(num_imgs), 10, 20, 40, 60, 80, max(num_imgs)))"] <- "image_number"

#print out numerical summary
imagesSummary
```
In the summary plot you can see if the amount of images affected the average, median, and standard deviation of shares.
